---
title: "Task1"
author: "Arman"
date: "10/14/2025"
format:
  html:
    embed-resources: true
editor: source
---

## Question 1 title

Question description

```{r}
# code for solving task
```


## Question N title

Question description

```{r}
# 1) Atomic types ----
# numeric
avg_temp_c  <- 23.7
radius_km   <- 6543.2
rating      <- 4.6
gdp_billion <- 312.8

# integer
n_moons       <- 2L
n_cities      <- 6L
inhabitants_m <- 1L

# logical
has_water    <- TRUE
is_inhabited <- TRUE

# character
planet_name <- "Aurelia"
sector      <- "Theta-9"
category    <- "Terra-class"

# S
cat(sprintf(
  "Planet %s in sector %s (%s): temp=%.1f°C, radius=%.1f km, rating=%.1f, GDP=%.1f bn; moons=%d, cities=%d, inhabitants~%dM; water=%s, inhabited=%s.\n",
  planet_name, sector, category, avg_temp_c, radius_km, rating, gdp_billion,
  n_moons, n_cities, inhabitants_m, has_water, is_inhabited
)) 
```
```{r}
# 2) Vectors
set.seed(42)

numeric_all <- c(avg_temp_c, radius_km, rating, gdp_billion,
                 n_moons, n_cities, inhabitants_m)
numeric_all

seq_vec <- seq(34, 100, by = 8)
seq_vec

logic_vec <- sample(c(TRUE, FALSE), size = 10, replace = TRUE)
logic_vec
## 3) Matrix

```{r}
res_matrix <- matrix(
  sample(-50:50, 9, replace = TRUE),
  nrow = 3, byrow = TRUE,
  dimnames = list(
    c("North_Ridge", "Equator_Bay", "South_Crater"),
    c("Energy", "Water", "Food")
  )
)
res_matrix

rowMeans(res_matrix)
colMeans(res_matrix)
```
## 4) List
```{r}
universe_list <- list(
  avg_temp_c = avg_temp_c,
  radius_km = radius_km,
  rating = rating,
  gdp_billion = gdp_billion,
  n_moons = n_moons,
  n_cities = n_cities,
  inhabitants_m = inhabitants_m,
  has_water = has_water,
  is_inhabited = is_inhabited,
  planet_name = planet_name,
  sector = sector,
  category = category,
  numeric_all = numeric_all,
  seq_vec = seq_vec,
  logic_vec = logic_vec,
  resources = res_matrix,
  description = "Aurelia is a temperate Terra-class world with six settlements."
)

str(universe_list)
```
## 5) Factor
planet_class <- factor(
  c("low", "medium", "high", "medium", "high", "low", "medium"),
  levels = c("low", "medium", "high"),
  ordered = TRUE
)
planet_class

table(planet_class)
## 6) Data frame
planet_df <- data.frame(
  city = c("Nova","Helios","Maris","Kron","Eden","Luna"),
  population_mil = c(1.2,0.8,2.3,1.7,0.5,0.9),
  has_water = c(TRUE, TRUE, TRUE, FALSE, TRUE, FALSE),
  rating = c(4.8,4.3,4.9,3.9,4.1,4.0),
  category = factor(c("high","medium","high","medium","low","medium"),
                    levels = c("low","medium","high"), ordered = TRUE)
)

names(planet_df) <- c("City","Population_Mil","Has_Water","Rating","Category")

str(planet_df)
summary(planet_df)

planet_df$Water_Rating_Ratio <- ifelse(
  planet_df$Has_Water, planet_df$Rating / planet_df$Population_Mil, NA
)

planet_df

